Fission product decay chain equilibria in \zirconia

### 2019-01-25 manuscript.tex
Add tellurium discussion

Using our generalised formula for $\mu_{O_2}$, we fix the temperature within the range of thermal phase-stabilisation (1500 K for tetragonal \zirconia) and calculate $\mu_{O_2}$ for many different values of $p_{O_2}$ between $10^{-35}$ and 10$^{0}$ atm, corresponding to oxygen deficient and oxygen rich environments, respectively ($p_{O_2}$ in air is approximately 0.2 atm). While the tetragonal phase will be stress-stabilised in practice, thermal-stabilisation in such models has been shown to qualitatively approximate the effect of stress-stabilisation, while allowing a wider range of dopant behaviours to be predicted \cite{Bell2016}. Equilibrium defect concentrations are then calculated at each $\mu_{O_2}$ and plotted against $p_{O_2}$ to produce a Brouwer diagram. Brouwer diagrams at extrinsic defect concentrations of $10^{-5}$ and $10^{-3}$ parts/fu (i.e. parts per \zirconia\ formula unit) were generated to examine low and high dopant concentrations, respectively. These two concentrations were examined because the amount of fission products present at a particular point in a fuel pellet depends on macroscopic parameters, including its position in the core and the time since the last shutdown, but also microscopic parameters such as the radial position in the pellet. These two concentrations were selected because $10^{-3}$ parts/fu is high enough to model an aggregation of iodine (such as at a crack tip), and $10^{-5}$ parts/fu was found to be the concentration below which iodine did not have a significant effect on defect equilibria.

@@

Brouwer diagrams for tellurium defects are shown in Figure \ref{figure:telluriumbrouwer}. At a tellurium defect concentration of $10^{-5}$ parts per f.u. \zirconia , the dominant defect was \ch{Te_{O}^{**}} over the oxygen pressure range of $10^{-28}$ to $10^{-7}$ atm. This is mainly charge compensated by intrinsic electronic defects at lower oxygen pressures, and zirconium vacancies at higher oxygen pressures. At higher oxygen pressures, tellurium occupies zirconium sites, producing the defect \ch{Te_{Zr}^{'''}}.

The defect volumes

@@

\begin{figure}
    \centering
    \includegraphics[height=11cm]{img/fissionyield.png}
    \caption{Plot of the percentage yield of nuclei with a given mass following a fission event. Range of masses corresponding to isotopes of Te, I, Xe and Cs are highlighted.}
    \label{fig:fissionyield}
\end{figure}

>>>>

Using our generalised formula for $\mu_{O_2}$, we fix the temperature within the range of thermal phase-stabilisation (1500 K for tetragonal \zirconia) and calculate $\mu_{O_2}$ for many different values of $p_{O_2}$ between $10^{-35}$ and 10$^{0}$ atm, corresponding to oxygen deficient and oxygen rich environments, respectively ($p_{O_2}$ in air is approximately 0.2 atm). While the tetragonal phase will be stress-stabilised in practice, thermal-stabilisation in such models has been shown to qualitatively approximate the effect of stress-stabilisation, while allowing a wider range of dopant behaviours to be predicted \cite{Bell2016}. Equilibrium defect concentrations are then calculated at each $\mu_{O_2}$ and plotted against $p_{O_2}$ to produce a Brouwer diagram. Brouwer diagrams at extrinsic defect concentrations of $10^{-5}$ and $10^{-3}$ parts/fu (i.e. parts per \zirconia\ formula unit) were generated to examine low and high dopant concentrations, respectively. These two concentrations were examined because the amount of fission products present at a particular point in a fuel pellet depends on macroscopic parameters, including its position in the core and the time since the last shutdown, but also microscopic parameters such as the radial position in the pellet. %These two concentrations were selected because $10^{-3}$ parts/fu is high enough to model an aggregation of iodine (such as at a crack tip), and $10^{-5}$ parts/fu was found to be the concentration below which iodine did not have a significant effect on defect equilibria.

@@

Brouwer diagrams for tellurium defects are shown in Figure \ref{figure:telluriumbrouwer}. At a tellurium defect concentration of $10^{-5}$ parts per f.u. \zirconia , the dominant defect was \ch{Te_{O}^{**}} over the oxygen pressure range of $10^{-28}$ to $10^{-7}$ atm. This is mainly charge compensated by intrinsic electronic defects at lower oxygen pressures, and zirconium vacancies at higher oxygen pressures. At higher oxygen pressures, tellurium preferentially occupies zirconium sites, producing the defect \ch{Te_{Zr}^{'''}}, this time charge compensated by positive hole defects. \\

Tellurium on an oxygen site adopts a 0 charge state, which while not obvious (it may be expected to act like oxygen, also a group 6 element), can be reasoned when considering chemical behaviour. Oxygen on its native site will readily adopt a charge state of -2. Tellurium on the other hand, exhibits more metallic behaviour, preferring to get oxidised itself. However, the \ch{V_{O}^{**}} site is highly reducing, and will counter the tellurium ion's proclivity to oxidise. The overall effect on tellurium in this case is to maintain its original electronic structure. Finally, when the tellurium concentration is increased to $10^{-3}$ parts/fu, greater occupancy on the zirconium site is predicted above oxygen pressures of $10^{-17}$ atm, forming the defect \ch{Te_{Zr}^{'''}}. This corresponds to a tellurium charge state of +1. While we would expect an even greater charge state for tellurium, in this case it is the dominant negative defect by an order of magnitude and is necessary to charge compensate positive hole and \ch{Te_{O}^{**}} defects. \\

In terms of defect volume, \ch{Te_{O}^{**}} has a slightly positive volume of 2.74 \r{A}$^{3}$. This is quite large for a positively charged defect (\ch{V_{O}^{**}} has a defect volume of -40.45 \r{A}$^{3}$), very likely due to the large size of the tellurium atom itself. When tellurium occupies the zirconium site, the resulting \ch{Te_{Zr}^{'''}} has a defect volume of 59.93 \r{A}$^{3}$. While this is lower than the defect volume of \ch{V_{Zr}^{''''}} (88.93 \r{A}$^{3}$)

@@

\begin{figure}
    \centering
    \includegraphics[height=11cm]{img/fissionyield.png}
    \caption{Plot of the percentage yield of nuclei with a given mass following a fission event. Range of masses corresponding to isotopes of Te, I, Xe and Cs are highlighted. Adapted from \cite{England1992OECDReport}.}
    \label{fig:fissionyield}
\end{figure}
